---
title: "Run_SBC"
output: html_document
date: "2025-02-17"
---

## packages and scripts


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages = c("brms","tidyverse","bayesplot","pracma","here",
             "patchwork","posterior","HDInterval","loo", "furrr", "SBC","future","mvtnorm")

do.call(pacman::p_load, as.list(packages))


source(here::here("Simulations","Psychometric","SBC","scripts.R"))
```


# test one dataset:
## make data and plot

```{r}
sim_data = generator_single(60,20)

sim_data_x = sim_data[[2]]

data.frame(sim_data_x) %>% ggplot(aes(x = x, y = binom_y))+geom_point()+facet_wrap(~S_id)
data.frame(sim_data_x) %>% ggplot(aes(x = x, y = RT))+geom_point()+facet_wrap(~S_id,scales = "free")
sim_data[[1]]

```

```{r}
#check RT's

# sim_data_x = sim_data[[2]]
# 
# data.frame(sim_data_x)  %>% 
#   # filter(RT > 10) %>%
#   ggplot(aes(x = RT)) + geom_histogram()
# 
# ids = unique(data.frame(sim_data_x)  %>% 
#   filter(RT > 10) %>% .$S_id)
# 
# sim_data[[1]][13:20]$rt_int[ids]
# sim_data[[1]][13:20]$rt_sd[ids]


```


## run one the single dataset:
```{r}
cmdstan_model <- cmdstanr::cmdstan_model(here::here("Simulations","Psychometric","SBC","stanmodels","subj_ndt.stan"))


fit = cmdstan_model$sample(data = sim_data_x,
                           iter_warmup = 250,
                           iter_sampling = 1000,
                           chains = 4,
                           # seed = 123,
                           refresh = 100,
                           max_treedepth = 10,
                           parallel_chains = 4,
                           adapt_delta = 0.95)

as_draws_df(fit$draws(c("gm","tau_u"))) %>% 
  select(-contains(".")) %>% 
  rename_with(~names(sim_data[[1]][1:12])) %>% 
  pivot_longer(everything()) %>% 
  mutate(draw = 1:n()) %>% 
  ggplot(aes(x = value))+
  geom_histogram(col = "black")+
  geom_vline(data = data.frame(sim_data[[1]][1:12]) %>% pivot_longer(everything()), aes(xintercept = value), col = "red", linetype = 2)+
  facet_wrap(~name, scales = "free")+
  theme_classic()


inner_join(
  as_draws_df(fit$draws(names(sim_data[[1]][13:20]))) %>% 
  select(-contains(".")) %>% 
  pivot_longer(everything()) %>% 
  mutate(draw = 1:n()) %>% 
  mutate(subject = as.numeric(sub(".*\\[(\\d+)\\].*", "\\1", name)),
         name = sub("\\[\\d+\\]", "", name)) %>% 
  group_by(name,subject) %>% 
  summarize(mean = mean(value),
            q5= quantile(value,0.05),
            q95= quantile(value,0.95))
  ,
  data.frame(sim_data[[1]][13:20]) %>% 
    mutate(subject = 1:n()) %>% 
    pivot_longer(-subject, values_to = "simulated_values")
  ) %>% 
  ggplot(aes(x = simulated_values, y = mean, ymin = q5, ymax = q95))+
  geom_pointrange()+
  facet_wrap(~name, scales = "free")+
  geom_abline()+
  theme_classic()


```


# run for all!
```{r}
n_sims <- 100 # Number of SBC iterations to run

generator <- SBC_generator_function(generator_single, N = 60, S = 10)

dataset <- generate_datasets(
  generator, 
  n_sims)

cmdstan_model <- cmdstanr::cmdstan_model(here::here("Simulations","Psychometric","SBC","stanmodels","subj_ndt.stan"))

backend <- SBC_backend_cmdstan_sample(
  cmdstan_model, iter_warmup = 1000, 
  iter_sampling = 2000,
  chains = 4, 
  max_treedepth = 12,
  parallel_chains = 4, 
  adapt_delta = 0.99)


library(future)
cores = 20
plan(multisession, workers = cores)

results <- compute_SBC(dataset,
                       ensure_num_ranks_divisor = 4,
                       keep_fits = F,
                       thin_ranks = 20,
                       #cores_per_fit = 10,
                       backend)


# save.image(here::here("Simulations","Psychometric","SBC","results","2000.RData"))

```

# load and removed badly or non-converged sims
```{r}
load(here::here("Simulations","Psychometric","SBC","results","2000.RData"))

results_psycho = results


# remove divergence and max tree depth:
sim_ids_to_keep <- results_psycho$backend_diagnostics %>%
  dplyr::filter(n_divergent == 0 & n_max_treedepth == 0) %>%
  dplyr::pull(sim_id)


sim_ids_to_exclude <- results_psycho$backend_diagnostics %>%
  dplyr::filter(n_divergent != 0 | n_max_treedepth != 0) %>%
  dplyr::pull(sim_id)


results_subset <- results_psycho[sim_ids_to_keep]
results_excluded <- results_psycho[sim_ids_to_exclude]

# also less than 400 effective samples

sim_ids_to_keep <- results_subset$default_diagnostics %>%
  dplyr::filter(min_ess_bulk  > 400)%>%
  dplyr::filter(min_ess_tail  > 400) %>%
  dplyr::pull(sim_id)


results_subset <- results_subset[sim_ids_to_keep]
```

# extraction and plotting

## diverenges vs no divergences (see if there are parameter regions that are bad)

```{r}
grouplevel = results_excluded$stats %>% 
  filter(grepl("^mu_", results_excluded$stats$variable) | grepl("^tau_", results_excluded$stats$variable)) %>% 
  mutate(variable = gsub("\\[\\d+\\]", "", variable))%>% select(sim_id, simulated_value,variable) %>% distinct() %>% 
  mutate(sim_id = as.factor(sim_id), excluded = T) 

grouplevel_inc = results_subset$stats %>% 
  filter(grepl("^mu_", results_subset$stats$variable) | grepl("^tau_", results_subset$stats$variable)) %>% 
  mutate(variable = gsub("\\[\\d+\\]", "", variable))%>% select(sim_id, simulated_value,variable) %>% distinct() %>% 
  mutate(sim_id = as.factor(sim_id), excluded = F)

rbind(grouplevel,grouplevel_inc)%>% ggplot(aes(x = simulated_value, fill = excluded))+geom_histogram(col = "black")+facet_wrap(~variable, scales = "free")
```


```{r}
subj_level = results_excluded$stats %>% 
  filter(!grepl("^mu_", results_excluded$stats$variable)& !grepl("^tau_", results_excluded$stats$variable)) %>% 
  mutate(variable = gsub("\\[\\d+\\]", "", variable))%>% select(sim_id, simulated_value,variable) %>% distinct() %>% 
  mutate(excluded = T)

subj_level_inc = results_subset$stats %>% 
  filter(!grepl("^mu_", results_subset$stats$variable)& !grepl("^tau_", results_subset$stats$variable)) %>% 
  mutate(variable = gsub("\\[\\d+\\]", "", variable))%>% select(sim_id, simulated_value,variable) %>% distinct()  %>% 
  mutate(excluded = F)


rbind(subj_level,subj_level_inc) %>% ggplot(aes(x = simulated_value, fill = excluded))+
  geom_histogram(col = "black")+facet_wrap(~variable, scales = "free")
```



```{r}
# Extract subject-level parameters from results2$stats
subject_level_params <- results_subset$stats %>% 
  filter(!grepl("^mu_", results_subset$stats$variable)& !grepl("^tau_", results_subset$stats$variable)) %>% 
  mutate(variable = gsub("\\[\\d+\\]", "", variable)) %>% 
  # filter(!sim_id %in% sims$sim_id) %>% 
  mutate()
# subject_level_params
plot_rank_hist(subject_level_params)
plot_sim_estimated(subject_level_params, alpha = 0.5)


```

# extract group level and plot:
```{r}
# Extract group-level parameters from results2$stats
group_level_params_psycho <- results_subset$stats %>% 
  filter(grepl("^mu_", results_subset$stats$variable) | grepl("^tau_", results_subset$stats$variable)) %>% 
  # filter(!sim_id %in% sims$sim_id) %>% 
  mutate()


histograms_psycho = plot_rank_hist(group_level_params_psycho)+facet_wrap(~variable, nrow = 2)+theme_minimal()
histograms_psycho
```


# only keep the group means and then also plot parameter recovery:

```{r}
means_group_level_params_psycho = group_level_params_psycho %>% filter(grepl("mu",variable))

histograms_psycho = plot_rank_hist(means_group_level_params_psycho)+facet_wrap(~variable, nrow = 1)+theme_minimal()
histograms_psycho

pp_recov_psycho = plot_sim_estimated(means_group_level_params_psycho, alpha = 0.5)+facet_wrap(~variable, nrow = 1, scales = "free")+theme_minimal()+
  theme(strip.text = element_blank())
```

# extract p-values for uniformity

```{r}
datas = data.frame()
parameters = unique(means_group_level_params_psycho$variable)

for (parameter in parameters){
  
  ranks = data.frame(means_group_level_params_psycho %>% filter(variable == parameter)) %>% .$rank
  
  observed <- table(factor(ranks, levels = 1:399))  # Force levels to include all 1:399
  
  # Calculate expected frequencies (uniform distribution)
  n_ranks <- 399  # Total number of unique ranks (1:399)
  total_count <- length(ranks)      # Total number of observations
  expected <- rep(total_count / n_ranks, n_ranks)
  
  qq = chisq.test(x = observed, p = expected / total_count)
  print(paste0("parameter = ", parameter, " p = ",qq$p.value))
  
  dd = data.frame(variable = parameter, pvalue = round(qq$p.value,2), chisq = round(qq$statistic,2))
  
  datas = rbind(dd,datas)
}
```

# combine this and plot it:

```{r}
histograms_psycho = histograms_psycho+
  geom_text(data = datas, aes(x = 75, y = 22, label = paste0("X^2 = ", chisq, "\np = ", pvalue)), size = 2)+
  facet_wrap(~variable, nrow = 1)+theme_minimal()

comb_psycho = (histograms_psycho + ggtitle("learning paradigm")) / pp_recov_psycho
comb_psycho
```

